### 2020-06-18 chapters/sota.tex
Some Background feedback applied

The authors split point outlier detection methods into uni and multivariate, as seen in Figure \ref{fig:outlier-taxonomy}. Furthermore, they split univariate point outlier detection methods into three broad categories: model-based, density-based and histogramming based \ref{fig:point-outlier-taxonomy}.

@@

Density-based univariate point outlier detection methods rely on the clustering of data points and neighbor counting. These methods consider a datum an outlier if it has less than $\alpha$ neighbors. Distance between two points is usually measured using Euclidean distance. Two points are said to be neighbors if they are close to each other --- \textit{i.e.} the distance between them is less than or equal to a given threshold $\alpha$.

Quoting the authors, histogramming univariate point outlier detection methods are \textit{"based on detecting the points whose removal from the univariate time series results in a histogram representation with lower error than the original"}.

@@

Ideally we want to maintain for each data point \textit{p\textsubscript{i}} the number of neighbors it has rather than the actual list of neighbors. The main advantage of this approach is the reduced memory footprint since only a count is stored rather than a list of pointers to all neighbors. The problem is that when a data point \textit{p\textsubscript{i}} expires, the count of each of \textit{p\textsubscript{i}}'s neighbors must be decremented by one but we are unable to reach \textit{p\textsubscript{i}}'s neighbors without a list of neighbor pointers. The solution to this problem is exploiting the previously presented \textit{predictability} property. The key idea is to predict the expiration of any data point \textit{p\textsubscript{i}} and pre-handle the impact it has on \textit{p\textsubscript{i}}'s neighbors. The authors introduce the concept of \textit{"lifetime neighbor counts"} or \textit{lt\_cnt}. The \textit{lt\_cnt} of a data point \textit{p\textsubscript{i}} corresponds to a sequence of \textit{predicted neighbor counts} --- \textit{i.e.} the number of \textit{predicted neighbors} \textit{p\textsubscript{i}} has in any future window where he exists. More precisely, each entry on \textit{pi.lt\_cnt} records the number of \textit{p\textsubscript{i}}'s neighbors that are known to survive in the corresponding future window. The authors provide the following example: consider a window \textit{W\textsubscript{i}} and a data point \textit{p\textsubscript{i}} with three neighbors, \textit{p\textsubscript{1}}, \textit{p\textsubscript{2}} and \textit{p\textsubscript{3}}. Applying the \textit{predictability} property makes it possible to compute the lifespan of all points. Assume \textit{p\textsubscript{1}} expires after \textit{W\textsubscript{i}}, \textit{p\textsubscript{2}} and \textit{p\textsubscript{3}} expire after \textit{W\textsubscript{i+1}} and \textit{p\textsubscript{i}} expires after \textit{W\textsubscript{i+2}}. This implies that at \textit{W\textsubscript{i}}, 
\[ \textit{p\textsubscript{i}.lt\_cnt} = \{\textit{W\textsubscript{i}}: 3, \textit{W\textsubscript{i+1}}: 2, \textit{W\textsubscript{i+2}}: 0\} \]
The lifetime neighbor counts (\textit{lt\_cnt}) carry enough information to compute density-based point outliers: for each data point \textit{p\textsubscript{i}} check if the first or current element of \textit{p\textsubscript{i}.lt\_cnt} is less than a user-defined threshold $\alpha$ to decide whether a data point is an outlier or not.

>>>>

The authors split point outlier detection methods into uni and multivariate, as seen in Figure \ref{fig:outlier-taxonomy}. Furthermore, they split univariate point outlier detection methods into three categories: model-based, density-based and histogram-based \ref{fig:point-outlier-taxonomy}.

@@

Density-based univariate point outlier detection methods rely on the clustering of data points and neighbor counting. These methods consider a datum an outlier if it has less than $\alpha$ neighbors. Distance between two points is usually measured using the Euclidean distance. Two points are said to be neighbors if they are close to each other --- \textit{i.e.} the distance between them is less than or equal to a given threshold $\alpha$.

Quoting the authors, histogram-based univariate point outlier detection methods are \textit{"based on detecting the points whose removal from the univariate time series results in a histogram representation with lower error than the original"}.

@@

Ideally we want to maintain for each data point \textit{p\textsubscript{i}}, the number of neighbors it has, rather than the actual list of neighbors. The main advantage of this approach is the reduced memory footprint since only a count is stored rather than a list of pointers to all neighbors. The problem is that when a data point \textit{p\textsubscript{i}} expires, the count of each of \textit{p\textsubscript{i}}'s neighbors must be decremented by one but we are unable to reach \textit{p\textsubscript{i}}'s neighbors without a list of neighbor pointers. The solution to this problem is exploiting the previously presented \textit{predictability} property. The key idea is to predict the expiration of any data point \textit{p\textsubscript{i}} and pre-handle the impact it has on \textit{p\textsubscript{i}}'s neighbors. The authors introduce the concept of \textit{"lifetime neighbor counts"} or \textit{lt\_cnt}. The \textit{lt\_cnt} of a data point \textit{p\textsubscript{i}} corresponds to a sequence of \textit{predicted neighbor counts} --- \textit{i.e.} the number of \textit{predicted neighbors} \textit{p\textsubscript{i}} has in any future window where he exists. More precisely, each entry on \textit{pi.lt\_cnt} records the number of \textit{p\textsubscript{i}}'s neighbors that are known to survive in the corresponding future window. The authors provide the following example: consider a window \textit{W\textsubscript{i}} and a data point \textit{p\textsubscript{i}} with three neighbors, \textit{p\textsubscript{1}}, \textit{p\textsubscript{2}} and \textit{p\textsubscript{3}}. Applying the \textit{predictability} property makes it possible to compute the lifespan of all points. Assume \textit{p\textsubscript{1}} expires after \textit{W\textsubscript{i}}, \textit{p\textsubscript{2}} and \textit{p\textsubscript{3}} expire after \textit{W\textsubscript{i+1}} and \textit{p\textsubscript{i}} expires after \textit{W\textsubscript{i+2}}. This implies that at \textit{W\textsubscript{i}}, 
\[ \textit{p\textsubscript{i}.lt\_cnt} = \{\textit{W\textsubscript{i}}: 3, \textit{W\textsubscript{i+1}}: 2, \textit{W\textsubscript{i+2}}: 0\} \]
The lifetime neighbor counts (\textit{lt\_cnt}) carry enough information to compute density-based point outliers: for each data point \textit{p\textsubscript{i}} check if the first or current element of \textit{p\textsubscript{i}.lt\_cnt} is less than a user-defined threshold $\alpha$ to decide whether a data point is an outlier or not.
